---
title: "Regex Actions"
output: html_notebook
---


# Import Dependencies


```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(stringi)
```

# Stringi String Documentation


```{r}
browseURL('http://www.gagolewski.com/software/stringi/manual/')
```


# A String & Regex Pattern to Match


```{r}
string <- c('fg44h4j', 'rghi', 'tr44e', 'IG4u')
pattern <- '4'
pattern2 <- 'i'
pattern3 <- 'g'
```

# Convert to Dataframe


```{r}
z <- data.frame(text = string, stringsAsFactors = FALSE)
z
```


# 4 Basic Regex Actions

1. Matching (logical/counting)
2. Subbing
3. Splitting
4. Extracting


## Matching: Logical


```{r}
stri_detect_regex(z$text, '4')
```


```{r}
stri_detect_regex(z$text, 'i')
```


### Ignore Case


```{r}
stri_detect_regex(z$text, '(?i)i')
```


## Matching: Counting


```{r}
stri_count_regex(z$text, '4')
```


```{r}
stri_count_regex(z$text, 'i')
```



### Ignore Case


```{r}
stri_count_regex(z$text, '(?i)i')
```


# Subbing


```{r}
stri_replace_all_regex(z$text, '4', '<<FOUR>>')
```





```{r}
stri_replace_all_regex(z$text, 'i', '<<AYE>>')
```




### Ignore Case


```{r}
stri_replace_all_regex(z$text, '(?i)i', '<<AYE>>')
```




# Splitting


```{r}
stri_split_regex(z$text, '4')
```



```{r}
stri_split_regex(z$text, 'g')
```




### Ignore Case


```{r}
stri_split_regex(z$text, '(?i)g')
```




### Splitting to Array


```{r}
y <- stri_split_regex(z$text, '(?i)g')
t(stri_list2matrix(y))
```




# Extracting


```{r}
stri_extract_all_regex(z$text, '(4)')
```


### Going Tall

```{r}
ex <- stri_extract_all_regex(z$text, '(4)')
data.frame(
	ind = rep(seq_along(ex), sapply(ex, length)),
	value = unlist(ex), 
	stringsAsFactors = FALSE
)
```




### Going Wide


```{r}
t(stri_list2matrix(ex))
```






```{r}
stri_extract_all_regex(z$text, '(g)')
```




### Ignore Case


```{r}
stri_extract_all_regex(z$text, '(?i)(g)')
```



# Exercises

## String Data


```{r}
wiki <- c("In theoretical computer science and formal language theory, a regular expression (sometimes called a rational expression)[1][2] is a sequence of characters that define a search pattern, mainly for use in pattern matching with strings, or string matching, i.e. \"find and replace\"-like operations.",
    "The concept arose in the 1950s, when the American mathematician Stephen Kleene formalized the description of a regular language, and came into common use with the Unix text processing utilities ed, an editor, and grep, a filter.",
    "In modern usage, \"regular expressions\" are often distinguished from the derived, but fundamentally distinct concepts of regex or regexp, which no longer describe a regular language.",
    "See below for details."
)
wiki
```



## Use `'wiki'` to make a 1 column dataframe with the column named wiki.


```{r}

```




## Which elements contain the expression "regular"?


```{r}

```


## Which elements contain the expression "regular expression"?


```{r}

```



## Get a count of the number of times the expression "regular" appears in each element.


```{r}

```




## Get a count of the number of times the expression "regular expression" appears in each element.


```{r}

```



## Get a count of the number of times the exact (match case) expression "the" appears in each element.


```{r}

```


## Get a count of the number of times the expression "the" appears in each element ignoring case.


```{r}

```



## Replace the expression "regular expression" with `"<<REGEX>>"`.


```{r}

```




## Replace the exact expression "the" with `"<<ARTICLE>>`.


```{r}

```






## Replace the expression "the" with `"<<ARTICLE>>` ignoring case.


```{r}

```




## Extract all the occurrences of the exact expression "the" from each element.


```{r}

```



## Extract all the occurrences of the expression "the" from each element ignoring case.


```{r}

```


## Split the text on the expression "the".


```{r}

```



## Split the text on the expression "the" and form a wide array.

```{r}

```



# Challenge

***Hint***: you'll need to use/read [the documentation](http://www.gagolewski.com/software/stringi/manual)







## Split on the first occurrence of the expression "the".


```{r}

```



## Split on the last occurrence of the expression "the".

```{r}

```
	
	
## Switch all lower case letter 'a' to 'e' and vice versa.

```{r}

```
	
