---
title: "Regex Basics Answer Key"
output: html_notebook
---


# Import Dependencies


```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(stringi, qdapRegex)
```

# Stringi String Documentation


```{r}
browseURL('http://www.gagolewski.com/software/stringi/manual/')
```


# A Regex Cheat Sheet


```
NAME                REGEX    WHAT IT DOES                                       
Exception           [^X]     Match everything except `X`                        
Dot                 .        Match any character                                
Case Insensitive    (?i)     Matches irregardless of case; `Foo` & `foO` matched
Digit               \\d      Match digits (i.e., [0-9])                         
Non-Digit           \\D      Match non-digits (i.e., [^0-9])                    
Word                \\w      Match words (i.e., [_a-zA-Z0-9])                   
Non-Word            \\W      Match non-words (i.e., [^_a-zA-Z0-9])              
Whitespace          \\s      Match whitespace (i.e., [ \\t\\r\\n\\f])           
Non-Whitespace      \\S      Match non-whitespace (i.e., [^ \\t\\r\\n\\f])      
Word Boundary       \\b      Match beginning/end of word                        
Non-Word Boundary   \\B      Match not beginning/end of word                    
0-1 (Greedy)        x?       Match 0-1 times greedy                             
0-1 (Lazy)          x??      Match 0-1 times lazy                               
>= 0 (Greedy)       x*       Match 0 or more times greedy                       
>= 0 (Lazy)         x*?      Match 0 or more times lazy                         
>= 1 (Greedy)       x+       Match 1 or more times greedy                       
>= 1 (Lazy)         x+?      Match 1 or more times lazy                         
Exactly N           x{4}     Match N times                                      
Min-Max             x{4,8}   Match min-max times                                
> N                 x{9,}    Match N or more times       
```

# Exercises

## String Data

```{r}
random <- c(
    "A download file from the http://example.com", 
    "Another url ftp://www.example.com",
    "And https://www.example.net",
    "@hadley (Dr. Wickham) I like #rstats for #ggplot2 work.",
    "Difference between #magrittr and #pipeR, both implement pipeline operators for #rstats: 
	http://renkun.me/r/2014/07/26/difference-between-magrittr-and-pipeR.html @timelyportfolio",
    "Is -2 an integer?", "-4.3 and 3.33 are not.",
    "123,456 is 0 alot -123456 more than -.2", "and 3456789123 fg for 345.",
    "There is $5.50 for me.", "that's 45.6% of the pizza", 
    "14% is $26 or $25.99",
    "Mr. Bean bought 2 tickets 2-613-213-4567",
    "43 Butter Rd, Brossard QC K0A 3P0 - 613 213 4567", 
    "The Rat Race, XX, 12345",
    "Ignore phone numbers(613)2134567",
    "Grab zips with dashes 12345-6789 or no space before12345-6789",  
    "Grab zips with spaces 12345 6789 or no space before12345 6789",
    "I like 1234567 dogs",
    " There is ( $5.50 ) for , me and you (NAME HERE). ", " that's [ 45.6% ] of! the pizza !", 
    "     14% is { $26  } or $25.99 ?", "Oh ;  here's colon : Yippee !",
    "I love chicken [unintelligible]!",
    "Me too Miss Jane! (laughter) It's so good.[interrupting]",
    "Yep it's awesome {reading}.", "Agreed Ms. Jones. {is so much fun}",
    "R uses 1:5 for 1, 2, 3, 4, 5.", 
    "At 3:00 we'll meet up , and we leave by 4:30:20",
    "We'll meet at 6:33 , bring $20,000.", "He ran it in :22.34"
)
random
```



## Use `'random'` to make a 1 column dataframe with the column named random.

```{r}
x <- data.frame(random =random, stringsAsFactors=FALSE)
x
```


## Extract all Twitter hash tags (#tag).

```{r}
stri_extract_all_regex(x$random, "#\\w+")
```


## Extract all Twitter user names (@name).

```{r}
stri_extract_all_regex(x$random, "@\\w+")
```


## Extract dollar amounts including dollars and decimals.

```{r}
stri_extract_all_regex(x$random, "\\$\\d[\\d,]*(\\.\\d+)*")
```


## Count the number of articles (the, a, an) in each string.  Remember to ignore case.  Don't count words that contain the articles (e.g., `"they"` contains the sub-string 'the'; don't count it).

```{r}
stri_count_regex(x$random, "(?i)\\b(the|a|an)\\b")
```



## Extract 5 or 9 (must contain a dash) digit zip codes.

```{r}
stri_extract_all_regex(x$random, "\\b\\d{5}(-\\d{4})?")
```


## Extract a name title (Mr., Ms., Mz., Miss, Dr., etc.) followed by a capitalized name (presumably last name).

```{r}
stri_extract_all_regex(x$random, "(Miss|((D|M)(s|r|rs|z)).)\\s+[A-Z][a-z']+")
```


## Replace double spaces with single spaces.

```{r}
stri_replace_all_regex(x$random, "\\s+", " ")
```


## Extract final punctuation mark. 

```{r}
stri_extract_all_regex(x$random, "[?.!]$")
```


## Which strings have a comma preceded by a space?  Remove space before commas.

```{r}
which(stri_count_regex(x$random, "\\s+,") > 0)
stri_replace_all_regex(x$random, "\\s+,", ",")
```


## Extract words starting with capital letters.

```{r}
stri_extract_all_regex(x$random, "\\b[A-Z]\\w+")
```


## Replace all parenthesis and the text in between with `'<<PLACEHOLDER>>'`.

```{r}
stri_replace_all_regex(x$random, "\\([^)]*?\\)", "<<PLACEHOLDER>>")
```


## Remove all non-vowel characters ignoring cases.

```{r}
stri_replace_all_regex(x$random, "(?i)[^aeiou]", "")
```


## Remove all URLs.

```{r}
stri_replace_all_regex(x$random, "(http[^ ]*)|(ftp[^ ]*)|(www\\.[^ ]*)", "")
```


## Tokenizing: 

1. Remove all characters except letters, apostrophes, & spaces.    
2. Remove leading and trailing spaces.    
3. Split each string into a vector of words (split on one or more spaces).      


```{r}
y <- stri_replace_all_regex(x$random, "[^A-Za-z' ]", " ")
y <- stri_replace_all_regex(y, "^\\s+|\\s+$", "")
stri_split_regex(y, "\\s+")
```


## Extract all numbers (including decimal points and negative signs).

***Note*** *that a number followed by a comma  (e.g., `"33,"` in `"We'll meet at 6:33, bring $20,000."`) might get captured.  We'll learn advanced techniques to handle this in a more advanced session.*

```{r}
stri_extract_all_regex(x$random, "-?\\d[\\d,]*(\\.\\d+)*")
```




