---
title: "Regex"
output: beamer_presentation
---

# My Regex Background

- 5 Years ago (replace human text coding)  
	- I mispronounce regex (!rejex)
	- Text analysis (extract patterns from transcripts)     
- 2 R regex based packages 

# What you will learn/get today?

1. Basic regex syntax
2. Resources for learning regex
3. Examples and growth exercises

# Agenda (~40 minutes)

1. Intro Regex (5 minutes)        
2. Action Tools: DS Prog. Functions (5 mins)    
3. Excercise \#1: Action Tools (8 mins)
4. Regex Basics (10 mins)
5. Exercise \#2: Basic Regex (remainder time)


# Regex Intro
5 minutes

# Additional Resources


- http://www.regular-expressions.info/tutorial.html    
- http://www.rexegg.com    
- https://www.debuggex.com

<!-- \vspace{2cm}\centering \includegraphics[scale=0.3]{regviz.png}-->

# What's it good for?

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
  \item Munging
  \item Quantifying text data
  \item Categorizing
  \item Validating 
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
    \begin{center}
     \includegraphics[width=1.0\textwidth]{chuck.jpg}
     \end{center}
\end{column}
\end{columns}


# What is regex?

> patterns of characters that match, or fail to match, sequences of characters in text. (Watt, 2005, p. 2)

# What is it for? 

## 4 Actions

1. Matching (logical/counting)
2. Subbing
3. Splitting
4. Extracting

# Action Tools
## Data Science Programing Language Functions

R & Python
5 Minutes

# Important Distinction

## Slashes for special characters:    
R - \\\\    
Python - \\


# Language Functions for 4 Actions

| Action | R: stringi  | Python: pandas |
|---------|-----|---------|
| Matching | `stri_detect_regex` | `.str.contains`     |
| Counting | `stri_count_regex`  | `.str.count`    |
| Subbing | `stri_replace_all_regex`  | `.str.replace`         |
| Splitting | `stri_split_regex` | `.str.split`      |
| Extracting | `stri_extract_all_regex` | `.str.extractall`   |

\vspace{3cm}
\tiny \emph{\textbf{Note}: There is a known bug in Pandas \texttt{.str.extractall} version 18.1}

# Regular Characters

Most characters, including all letters and digits, are regular expressions that match themselves

# Exercise \#1: Action Tools 

8 minutes

```{r, echo=FALSE, comment = NA}
stopsign <- c("         ______________", "       /.--------------.\\", "      //                \\\\", 
    "     //                  \\\\", "    || .-..----. .-. .--. ||", 
    "    ||( ( '-..-'|.-.||.-.|||", "    || \\ \\  ||  || ||||_||||", 
    "    ||._) ) ||  \\'-'/||-' ||", "     \\\\'-'  `'   `-' `'  //", 
    "      \\\\                //", "       \\\\______________//", 
    "        '--------------'", "              |_|_", "       ____ _/ _)_)", 
    "           '  | (_)", "        .--'\"\\| ()", "      mx      | |", 
    "              | |", "              |_|")


        
#cat(paste(stopsign, collapse="\n"))       
```

# Regex Basics 

10 minutes

# Meta-Characters

| Metacharacter | Meaning           |
|--------------|-------------------|
| ^ | Beginning of string           |
| $ | End of string                 |
| . | Any character                 |
| `*` | Match 0 or more times         |
| + | Match 1 or more times         |
| ? | Match 0 or 1 times            |
| `|` | Or                            |
| ( ) | Group                       |
| [ ] | Set of characters           |
| { } | Repetition modifier          |
| \ | Escape quote or metacharacter |


## Escape 'em

Escape these (usually with a `\`)

# Mental Excercise

Write a regular expression to match $1.00


# Mental Excercise

Write a regular expression to match $1.00

\vspace{1in}

**\\$1\\.00**

# Character Classes

- []
- special
	- 0-9
	- A-Z
	- a-z
- meta-characters (you have no power here; only `]`, `\`, `^`, `-`)
- no ordering; e.g., [xy] == [yx]

\centering
\vspace{2cm}
`[A-Za-z' ]`

	
# Character Class: Negation

`[^expressions_here]` - Everything except...

\centering
\vspace{5cm}
`[^0-9.-]`

# Short Hand Character Class

\footnotesize

| Regex          | Name           | Action   |
|----------------|----------------|----------|
| `\d`            | Digit          | Match     digits                                    |
| `\D`            | Non-Digit      | Match     non-digits                             |
| `\w`            | Word           | Match     words                   |
| `\W`            | Non-Word       | Match     non-words                |
| `\s`            | Whitespace     | Match     whitespace             |
| `\S`            | Non-Whitespace | Match     non-whitespace                 |  

\**Most Useful: `\d` & `\s`*

\centering
\vspace{2cm}
`\w` = [A-Za-z0-9_]

# Quantifiers

\footnotesize

+---------+---------------+------------------------------+
|  Regex  |     Name      |         What it Does         |
+=========+===============+==============================+
|   `x?`    | 0-1 (Greedy)  |    Match 0-1 times greedy    |
+---------+---------------+------------------------------+
|  `x??`   |  0-1 (Lazy)   |     Match 0-1 times lazy     |
+---------+---------------+------------------------------+
|   `x*`    | >= 0 (Greedy) | Match 0 or more times greedy |
+---------+---------------+------------------------------+
|   `x*?`   |  >= 0 (Lazy)  |  Match 0 or more times lazy  |
+---------+---------------+------------------------------+
|   `x+`    | >= 1 (Greedy) | Match 1 or more times greedy |
+---------+---------------+------------------------------+
|   `x+?`   |  >= 1 (Lazy)  |  Match 1 or more times lazy  |
+---------+---------------+------------------------------+
|  `x{4}`   |   Exactly N   |        Match N times         |
+---------+---------------+------------------------------+
| `x{4,8}`  |    Min-Max    |     Match min-max times      |
+---------+---------------+------------------------------+
|  `x{9,}`  |      > N      |    Match N or more times     |
+---------+---------------+------------------------------+

Use with single chars or [] or ()

# Quantifiers: Greedy vs. Lazy

```{r}
gsub('\\(.*\\)', "<<OUT>>", "Look at (A) and then (B).")
gsub('\\(.*?\\)', "<<OUT>>", "Look at (A) and then (B).")
```

# Quantifiers: Greedy vs. Lazy

```{r}
gsub('\\([^)]*\\)', "<<OUT>>", "Look at (A) and then (B).")
```


# Boolean Or

`|` - pipe   

\centering
\vspace{2cm}

`a|e`

# Grouping

`()` - Group

- Order matters (unlike character classes `[]`)
- Useful with quantifiers    

\centering
\vspace{2cm}

`gr(a|e)y`    
`(cat|dog|fish)`    
`(?i)(hey jude ){2,}`
 



# Anchors & Boundaries

- ^ - begining     
- $ - end     
- `\b` - word boundary 

\vspace{.5cm}
\*[raw string (r'\\b') for Python](http://stackoverflow.com/a/3995242/1000343)


\centering
\vspace{2cm}

`^\w+`     
`\w+[.?!]+$`    
`\bread`    
`\bread\b`     





